---
title: "Assignment1_ML"
author: "Tanu Sreedharan"
date: "1/26/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library (viridis)
knitr::opts_chunk$set(
echo = TRUE,
warning = FALSE,
fig.width = 8, 
 fig.height = 6,
 out.width = "90%"
)
options(
 ggplot2.continuous.colour = "viridis",
 ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)

```

```{r}
bc_data = read_csv(file = "Data/dataR2.csv") %>% 
  janitor::clean_names()
```

#### A table providing summaries of the quantitative features of the dataset. Summaries includes the mean, median, minimum value, and maximum value.

```{r}
bc_data %>% 
select(-classification) %>%   
summary(bc_data) %>% 
  knitr::kable(col.names = c("Age",
                            "BMI",
                            "Glucose",
                            "Insulin",
                            "Homa",
                            "Leptin",
                            "Adiponectin",
                            "Resistin",
                            "mcp_1"))
```


##### Recoded BMI into the WHO-defined categories.
 
 
```{r}
Breast_Cancer=
  bc_data %>% 
  mutate(BMI_CAT=
           case_when(bmi < 16.5 ~ "Severly Underweight",
                     bmi >= 16.5 & bmi < 18.5 ~ "Underweight",
                     bmi >= 18.5 & bmi <= 24.9 ~ "Normal Weight",
                     bmi >= 25 & bmi <= 29.9 ~ "Overweight",
                     bmi >= 30 & bmi <= 34.9 ~ "Obesity Class I",
                     bmi >= 35 & bmi <= 39.9 ~ "Obesity Class II",
                     bmi >= 40 ~ "Obesity Class III"))
```


#### 3.Created a bar chart showing the proportion of breast cancer cases and controls within each BMI category


```{r}
Chart_Data =
Breast_Cancer %>% 
  mutate(classification = recode( classification, "1" = "Control", "2" = "Case"))
```

```{r}
Chart_Data %>% 
  ggplot(aes(x=BMI_CAT, fill=classification)) +
  geom_bar() + 
 theme_minimal() +
  labs(
    title = "Proportion of breast cancer cases and controls by BMI Category" ,
      x= "BMI Category",
      y = "Count"
  ) 
```


#### Logistic regression model using breast cancer classification as the outcome and glucose, HOMA, leptin, BMI (continuous) and age as the independent variables. 


### Filled in the beta estimate and 95% confidence interval associated with a 1-unit change in HOMA.


```{r}
bc_logmodel=lm(classification ~ glucose + homa + leptin + bmi + age, data = Breast_Cancer, family = binomial())

bc_logmodel %>% 
  broom::tidy() %>% 
  filter(term == "homa") %>%
  mutate(OR = exp(estimate),
         CI_UL= exp(estimate + 1.96 * std.error),
         CI_LL= exp(estimate - 1.96 * std.error)) %>%
  select(term, estimate, OR, CI_LL, CI_UL) %>% 
  knitr::kable(col.names = c("Term",
                             "Beta Estimate",
                             "Odds Ratio",
                             "95%CI Lower Limit",
                             "95%CI Upper Limit"),
                               digits = 3)

# At a 5% level of significance, the odds of homa among the cases is 1.005 (95%CI 0.97, 1.04) times the odds of homa among the controls.

```


#### Linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables. 

### Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in age.


```{r}
bc_linmodel = glm(insulin ~ bmi + age + glucose, data = Breast_Cancer)

bc_linmodel %>% 
  broom::tidy() %>% 
  filter(term == "age") %>%
  mutate(CI_UL= exp(estimate + 1.96 * std.error),
         CI_LL= exp(estimate - 1.96 * std.error)) %>%
  select(term, estimate, CI_LL, CI_UL) %>% 
  knitr::kable(col.names = c("Term",
                             "Beta Estimate",
                             "95%CI Lower Limit",
                             "95%CI Upper Limit"),
                               digits = 3)    

# At a 5% level of significance, insulin intake decreases on average by 0.05 (95% CI 0.86, 1.05) for every one unit increase in age.
```
